---
title: "Project-AlGhammari-Schmidt"
author: "Bernd Schmidt , Osamah Al-Ghammari"
date: "November 14, 2017"
output: pdf_document
---

```{r setup, message=FALSE, results='hide'}
require(plyr)
require(corrplot)
require(caret)
require(doMC)
registerDoMC(3)
require(zoo)
```

# Project MyoHand
Detecting gestures from EMG data provided by the MYO Armband.

```{r read data}
filedir <- '../dat'

filenames <- list.files(filedir, full.names = T, pattern="*.csv")

gestures <- list()
gestures$emg0 <- ldply(filenames[1], read.table, sep=',', fill = T, col.names = c('gesture', 'person', 'sample', paste('emg', 1:250, sep='')))
gestures$emg1 <- ldply(filenames[2], read.table, sep=',', fill = T, col.names = c('gesture', 'person', 'sample', paste('emg', 1:250, sep='')))
gestures$emg2 <- ldply(filenames[3], read.table, sep=',', fill = T, col.names = c('gesture', 'person', 'sample', paste('emg', 1:250, sep='')))
gestures$emg3 <- ldply(filenames[4], read.table, sep=',', fill = T, col.names = c('gesture', 'person', 'sample', paste('emg', 1:250, sep='')))
gestures$emg4 <- ldply(filenames[5], read.table, sep=',', fill = T, col.names = c('gesture', 'person', 'sample', paste('emg', 1:250, sep='')))
gestures$emg5 <- ldply(filenames[6], read.table, sep=',', fill = T, col.names = c('gesture', 'person', 'sample', paste('emg', 1:250, sep='')))
gestures$emg6 <- ldply(filenames[7], read.table, sep=',', fill = T, col.names = c('gesture', 'person', 'sample', paste('emg', 1:250, sep='')))
gestures$emg7 <- ldply(filenames[8], read.table, sep=',', fill = T, col.names = c('gesture', 'person', 'sample', paste('emg', 1:250, sep='')))
```

## Calculating the variance threshhold
To detect the start and the end of each gesture, the variance over a specific window is calculated.
If the value is greater than the threshold, the signal will be cut.

### Magnitude over all 8 EMG values
For this purpose we will utilize the magnitude of all eight EMG values

```{r calculate magnitude}
gestures$mag <- as.data.frame(mapply(function(emg0, emg1, emg2, emg3, emg4, emg5, emg6, emg7) {
  m <- sqrt(emg0^2 + emg1^2 + emg2^2 + emg3^2 + emg4^2 + emg5^2 + emg6^2 + emg7^2)
  m
}, gestures$emg0[,4:253], gestures$emg1[,4:253], gestures$emg2[,4:253], gestures$emg3[,4:253], gestures$emg4[,4:253], gestures$emg5[,4:253], gestures$emg6[,4:253], gestures$emg7[,4:253]))
matplot(t(gestures$mag[1:10,]), type='l')
```

### Calculate variance using the 25% quantile

```{r sd of each sample}
gestures$q25 <- as.data.frame(apply(gestures$mag, 1 ,function(m) {
  q25 <- quantile(m, na.rm = T)[[3]]
  q25
}))
```

### Stripping the data for all values below this quanitle

```{r strip values}
stepwidth <- 1/250
gestures$new_mag <- data.frame(matrix(NA, ncol = 250))
gestures$data <- data.frame(matrix(NA, ncol = 2003))
for (i in 1:dim(gestures$mag)[1]) {
  # 1 remove NAs
  
  # 2 pre filter
  mag <- gestures$mag[i,]
  #rollapply(mag, 11, median, na.rm = T)
  # 3 cut
  start <- which.max(mag > gestures$q25[i,])
  stop <- length(mag) - which.max(rev(mag) > gestures$q25[i,])
  mag <- mag[start:stop]
  start <- start + 3
  stop <- stop + 3
  emg0 <- gestures$emg0[i,start:stop]
  emg1 <- gestures$emg1[i,start:stop]
  emg2 <- gestures$emg2[i,start:stop]
  emg3 <- gestures$emg3[i,start:stop]
  emg4 <- gestures$emg4[i,start:stop]
  emg5 <- gestures$emg5[i,start:stop]
  emg6 <- gestures$emg6[i,start:stop]
  emg7 <- gestures$emg7[i,start:stop]
  # remove NAs
  emg0 <- emg0[,!is.na(emg0)]
  emg1 <- emg1[,!is.na(emg1)]
  emg2 <- emg2[,!is.na(emg2)]
  emg3 <- emg3[,!is.na(emg3)]
  emg4 <- emg4[,!is.na(emg4)]
  emg5 <- emg5[,!is.na(emg5)]
  emg6 <- emg6[,!is.na(emg6)]
  emg7 <- emg7[,!is.na(emg7)]
  # 4 approx
  mag_approx <- approx(x = seq(0,1,1/(length(mag)-1)), y = mag, xout = seq(0,1,stepwidth), method = 'linear')$y[1:250]
  
  emg0_approx <- approx(x = seq(0,1,1/(length(emg0)-1)), y = emg0, xout = seq(0,1,stepwidth), method = 'linear')$y[1:250]
  emg1_approx <- approx(x = seq(0,1,1/(length(emg1)-1)), y = emg1, xout = seq(0,1,stepwidth), method = 'linear')$y[1:250]
  emg2_approx <- approx(x = seq(0,1,1/(length(emg2)-1)), y = emg2, xout = seq(0,1,stepwidth), method = 'linear')$y[1:250]
  emg3_approx <- approx(x = seq(0,1,1/(length(emg3)-1)), y = emg3, xout = seq(0,1,stepwidth), method = 'linear')$y[1:250]
  emg4_approx <- approx(x = seq(0,1,1/(length(emg4)-1)), y = emg4, xout = seq(0,1,stepwidth), method = 'linear')$y[1:250]
  emg5_approx <- approx(x = seq(0,1,1/(length(emg5)-1)), y = emg5, xout = seq(0,1,stepwidth), method = 'linear')$y[1:250]
  emg6_approx <- approx(x = seq(0,1,1/(length(emg6)-1)), y = emg6, xout = seq(0,1,stepwidth), method = 'linear')$y[1:250]
  emg7_approx <- approx(x = seq(0,1,1/(length(emg7)-1)), y = emg7, xout = seq(0,1,stepwidth), method = 'linear')$y[1:250]
  # 5 filter
  
  gestures$new_mag[i,] <- as.data.frame(t(mag_approx))
  gestures$emg0[i,4:253] <- as.data.frame(t(emg0_approx))
  gestures$emg1[i,4:253] <- as.data.frame(t(emg1_approx))
  gestures$emg2[i,4:253] <- as.data.frame(t(emg2_approx))
  gestures$emg3[i,4:253] <- as.data.frame(t(emg3_approx))
  gestures$emg4[i,4:253] <- as.data.frame(t(emg4_approx))
  gestures$emg5[i,4:253] <- as.data.frame(t(emg5_approx))
  gestures$emg6[i,4:253] <- as.data.frame(t(emg6_approx))
  gestures$emg7[i,4:253] <- as.data.frame(t(emg7_approx))
  
  gestures$data[i,1:253] <- gestures$emg0[i,]
  gestures$data[i,254:503] <- gestures$emg1[i,4:253]
  gestures$data[i,504:753] <- gestures$emg2[i,4:253]
  gestures$data[i,754:1003] <- gestures$emg3[i,4:253]
  gestures$data[i,1004:1253] <- gestures$emg4[i,4:253]
  gestures$data[i,1254:1503] <- gestures$emg5[i,4:253]
  gestures$data[i,1504:1753] <- gestures$emg6[i,4:253]
  gestures$data[i,1754:2003] <- gestures$emg7[i,4:253]
}
matplot(t(gestures$new_mag[1:10,]), type='l')
matplot(t(gestures$emg0[1:10,-(1:3)]), type='l')
```
## Data Validation and Optimization

### Feature Correlation

```{r feature correlation}
# feature correlation as plot
corrplot(cor(gestures$data[,4:100]), tl.cex = 0.3) # addgrid.col = NA
# remove correlated variable using ?findCorrelation
foundCorIndexes <- findCorrelation(cor(gestures$data))
#foundCorIndexes
corrplot(cor(gestures$data[,-foundCorIndexes]), tl.cex = 0.3)
# remove the features from the data
gestures$data <- gestures$data[,-foundCorIndexes]
```

## Data Partitioning

```{r data_partitioning}
# split into training and test data
set.seed(1704)
indexes_train <- createDataPartition(gestures$emg0[,1], p=0.75, list = F)
indexes_test <- (1:nrow(gestures$data))[-indexes_train]

training <- gestures$data[indexes_train,]
training_gest <- gestures$emg0[indexes_train,1]
testing <- gestures$data[indexes_test,]
testing_gest <- gestures$emg0[indexes_test,1]
```

## Model Training

```{r initialize models list}
models <- list()
```

```{r specify train control}
trControl <- trainControl(
    method = 'repeatedcv', # none, cv, repeatedcv, LOOCV, ...
    number = 10, # nr of CV partitions
    repeats = 20, # nr of partitioning repetitions
    returnData = F, 
    # classProbs = T, # enable computation of class probabilities?
    # summaryFunction = twoClassSummary, # use when classifying two classes 
    returnResamp = 'final', # return CV partition results for best model
    allowParallel = T
)

#trControl <- trainControl(
#                 method = 'LOOCV',
#                 preProcOptions = list(thresh = 0.9),
#                 returnResamp = 'final',
#                 returnData = F,
#                 savePredictions = T,
#                 allowParallel = T
#               )
```

### KNN

```{r model training knn}
models <- list()
models$knn <- train(training,
               factor(training_gest),
               method = 'knn',
               preProcess = c('center', 'scale', 'pca'),
               metric = 'Kappa',
               trControl = trControl
               )
models$knn
varImp(models$knn)
```

```{r predict_test_data knn}
predicted <- predict(models$knn, newdata = testing)

# to ensure, that also when one level is not predicted, the results can be displayed
u = union(predicted, testing_gest)
t = table(factor(predicted, u), factor(testing_gest, u))
conf <- confusionMatrix(t)

levelplot(sweep(conf$table, MARGIN = 2, STATS = colSums(conf$table), FUN = `/`), col.regions = gray(100:0/100))
```

### LDA
To compare the results, now a *lda* model with the same parameters is trained.

```{r model training lda}
models$lda <- train(training,
               factor(training_gest),
               method = 'lda',
               preProcess = c('center', 'scale', 'pca'),
               metric = 'Kappa',
               trControl = trControl
               )
models$lda
```

```{r predict_test_data lda}
predicted <- predict(models$lda, newdata = testing)

# to ensure, that also when one level is not predicted, the results can be displayed
u = union(predicted, testing_gest)
t = table(factor(predicted, u), factor(testing_gest, u))
conf <- confusionMatrix(t)

levelplot(sweep(conf$table, MARGIN = 2, STATS = colSums(conf$table), FUN = `/`), col.regions = gray(100:0/100))
```

### LDA2

```{r model training lda2}
models$lda2 <- train(training,
               factor(training_gest),
               method = 'lda2',
               preProcess = c('center', 'scale', 'pca'),
               metric = 'Kappa',
               trControl = trControl
               )
models$lda2
```

```{r predict_test_data lda2}
predicted <- predict(models$lda2, newdata = testing)

# to ensure, that also when one level is not predicted, the results can be displayed
u = union(predicted, testing$pers)
t = table(factor(predicted, u), factor(testing_gest, u))
conf <- confusionMatrix(t)

levelplot(sweep(conf$table, MARGIN = 2, STATS = colSums(conf$table), FUN = `/`), col.regions = gray(100:0/100))
```

### SVM

```{r train model svm}
train_model <- function(method, tuneGrid=NULL) {
  train(x = training, # in real life apps only use train data here!
        y = training_gest, # in real life apps only use train data here!
        method = method, 
        metric = 'Kappa', 
        tuneGrid = tuneGrid,
        trControl = trControl
  )
}
models$svmLinear <- train_model('svmLinear', tuneGrid = expand.grid(C=3**(-5:5)))
models$svmRadial <- train_model('svmRadial', tuneGrid = expand.grid(C=3**(-5:5), sigma=3**(-5:5)))
```

```{r display model svm}
print(plot(models$svmLinear, scales=list(x=list(log=3))))
print(plot(models$svmRadial, scales=list(x=list(log=3))))
```

```{r predict_test_data svm linear}
predicted <- predict(models$svmLinear, newdata = testing)

# to ensure, that also when one level is not predicted, the results can be displayed
u = union(predicted, testing$pers)
t = table(factor(predicted, u), factor(testing_gest, u))
conf <- confusionMatrix(t)

levelplot(sweep(conf$table, MARGIN = 2, STATS = colSums(conf$table), FUN = `/`), col.regions = gray(100:0/100))
```

```{r predict_test_data svm radial}
predicted <- predict(models$svmRadial, newdata = testing)

# to ensure, that also when one level is not predicted, the results can be displayed
u = union(predicted, testing$pers)
t = table(factor(predicted, u), factor(testing_gest, u))
conf <- confusionMatrix(t)

levelplot(sweep(conf$table, MARGIN = 2, STATS = colSums(conf$table), FUN = `/`), col.regions = gray(100:0/100))
```

## Result Comparison

```{r result comparison}
# save models to file to ensure that the results were not lost
saveRDS(object = models, file = "project_models.RDS")
results <- resamples(models)
summary(results)
bwplot(results)
```